---
title: "financial data analysis"
author: "iris mancini"
format: html
editor: visual
---

```{r, message = FALSE}

here::i_am("dplyr-102.Rproj")
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(vroom)

```

## Data loading

```{r, message = FALSE}
clients <- vroom(here("data", "client.csv")) 
#it gives you the nb of columns and rows, if they are not correct the limiter is wrong
accounts <- vroom(here("data", "account.csv"))
disp <- vroom(here("data", "disp.csv"))
```

## Joining data tables 

A collection of joining functions of the form "\*\_join".

### Inner join : does by default what is called the natural join

we want to join the clients with their accounts :

```{r}
#disposition and clients : 

da <- inner_join(accounts, disp) #displays with what variable the join has been done, here it is account_id

```

```{r}
cda <- inner_join(clients, da, by = join_by(client_id), suffix = c("_home", "_bank"))
  
#dans cda on a deux nouvelles colonnes qui se créent : district_id.x et district_id.y, chacune des colonnes correspond à un district tiré d'une table différente. home va être pris dans clients et bank pour da. 

```

### Home district versus bank district : 

```{r}
home_diff_bank <- cda |> filter(district_id_home != district_id_bank) |> nrow()
```

We observe that `r home_diff_bank` clients have a bank in a different district than where they live.

```{r}
cda |> 
  group_by(account_id) |>
  distinct(district_id_home) |> #garde que ceux qui ont les mêmes account et pas mm district
  summarise(N =n()) |> 
  filter(N>1)
  
```

All the people who have the same account live in the same district.

### Gender gap : 

```{r}
clients |> 
  ggplot(aes(x=gender)) + geom_bar() + coord_flip() 


```

```{r}
cda |> 
  count(gender, type) |>
  group_by(gender) |>
  mutate(freq=n/sum(n)) |>
  select(-n) |>
  pivot_wider(values_from = freq, names_from = type)
```

```{r}
cda |>
  ggplot(aes(y=gender, fill = type)) + geom_bar(position = "fill")
```

```{r}
chisq.test(cda |> pull(gender), 
           cda |> pull(type))
#H0 not rejected, no dependecy on the gender to type (owner or disponent)
```
